---
title: "Blast analysis of species removal"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Given a reference tree we ran the final step of r2t using the '--remove_species_mapping' parameter increasing at each iteration the number of species not to be considered to reconstruct our sequence. Therefore at each iteration the closest neighbour is more distantly related to the species we are trying to map. We then use our obtained sequences and blast them against their original OGs containing all the sequences including the species we are trying to map.  


```{r echo=FALSE}
library(ggplot2)
library(plyr)
#library(dplyr)
library(reshape2)
library(base)
library(grid)
library(gtable)
library(viridis)
theme_set(
  theme_light() + theme(legend.position = "top")
  )
# source("/Volumes/Recovery/Dropbox/dessimoz/research/opt/R-pLIBS/ggplot_pLIB.R")
source("https://raw.githubusercontent.com/teunbrand/ggnomics/standalone-facet_nested/R/facet_nested.R")

## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    #datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}

OverlappingStripLabels = function(plot) {

  # Get the ggplot grob
  pg = ggplotGrob(plot)

  ### Collect some information about the strips from the plot
  # Get a list of strips
  stripr = lapply(grep("strip-r", pg$layout$name), function(x) {pg$grobs[[x]]})

  stript = lapply(grep("strip-t", pg$layout$name), function(x) {pg$grobs[[x]]})

  # Number of strips
  NumberOfStripsr = sum(grepl(pattern = "strip-r", pg$layout$name))
  NumberOfStripst = sum(grepl(pattern = "strip-t", pg$layout$name))

  # Number of columns
  NumberOfCols = length(stripr[[1]])
  NumberOfRows = length(stript[[1]])

  # Panel spacing
  plot_theme <- function(p) {
    plyr::defaults(p$theme, theme_get())
  }
  PanelSpacing = plot_theme(plot)$panel.spacing

  # Map the boundaries of the new strips
  Nlabelr = vector("list", NumberOfCols)
  mapr = vector("list", NumberOfCols)
  for(i in 1:NumberOfCols) {

    for(j in 1:NumberOfStripsr) {
      Nlabelr[[i]][j] = getGrob(grid.force(stripr[[j]]$grobs[[i]]), gPath("GRID.text"), grep = TRUE)$label
    }

    mapr[[i]][1] = TRUE
    for(j in 2:NumberOfStripsr) {
      mapr[[i]][j] = as.character(Nlabelr[[i]][j]) != as.character(Nlabelr[[i]][j-1])#Nlabelr[[i]][j] != Nlabelr[[i]][j-1]
    }
  }

  # Map the boundaries of the new strips
  Nlabelt = vector("list", NumberOfRows)
  mapt = vector("list", NumberOfRows)
  for(i in 1:NumberOfRows) {

    for(j in 1:NumberOfStripst) {
      Nlabelt[[i]][j] = getGrob(grid.force(stript[[j]]$grobs[[i]]), gPath("GRID.text"), grep = TRUE)$label
    }

    mapt[[i]][1] = TRUE
    for(j in 2:NumberOfStripst) {
      mapt[[i]][j] = as.character(Nlabelt[[i]][j]) != as.character(Nlabelt[[i]][j-1])#Nlabelt[[i]][j] != Nlabelt[[i]][j-1]
    }
  }


  ## Construct gtable to contain the new strip
  newStripr  = gtable(heights = unit.c(rep(unit.c(unit(1, "null"), PanelSpacing), NumberOfStripsr-1), unit(1, "null")), 
                     widths = stripr[[1]]$widths)
  ## Populate the gtable  
  seqTop = list()
  for(i in NumberOfCols:1) {  
    Top = which(mapr[[i]] == TRUE)
    seqTop[[i]] = if(i == NumberOfCols) 2*Top - 1 else  sort(unique(c(seqTop[[i+1]], 2*Top - 1)))  
    seqBottom = c(seqTop[[i]][-1] -2, (2*NumberOfStripsr-1))
    newStripr = gtable_add_grob(newStripr, lapply(stripr[(seqTop[[i]]+1)/2], function(x) x[[1]][[i]]), l = i, t = seqTop[[i]], b = seqBottom)
  }

  mapt <- mapt[NumberOfRows:1]
  Nlabelt <- Nlabelt[NumberOfRows:1]
  ## Do the same for top facets
  newStript  = gtable(heights = stript[[1]]$heights,
                      widths = unit.c(rep(unit.c(unit(1, "null"), PanelSpacing), NumberOfStripst-1), unit(1, "null")))
  seqTop = list()
  for(i in NumberOfRows:1) {  
    Top = which(mapt[[i]] == TRUE)
    seqTop[[i]] = if(i == NumberOfRows) 2*Top - 1 else  sort(unique(c(seqTop[[i+1]], 2*Top - 1)))  
    seqBottom = c(seqTop[[i]][-1] -2, (2*NumberOfStripst-1))
    # newStript = gtable_add_grob(newStript, lapply(stript[(seqTop[[i]]+1)/2], function(x) x[[1]][[i]]), l = i, t = seqTop[[i]], b = seqBottom)
    newStript = gtable_add_grob(newStript, lapply(stript[(seqTop[[i]]+1)/2], function(x) x[[1]][[(NumberOfRows:1)[i]]]), t = (NumberOfRows:1)[i], l = seqTop[[i]], r = seqBottom)
  }

  ## Put the strip into the plot
  # Get the locations of the original strips
  posr = subset(pg$layout, grepl("strip-r", pg$layout$name), t:r)
  post = subset(pg$layout, grepl("strip-t", pg$layout$name), t:r)

  ## Use these to position the new strip
  pgNew = gtable_add_grob(pg, newStripr, t = min(posr$t), l = unique(posr$l), b = max(posr$b))
  pgNew = gtable_add_grob(pgNew, newStript, l = min(post$l), r = max(post$r), t=unique(post$t))
  grid.draw(pgNew)

  return(pgNew)
}
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

```{r}
mouse <- read.csv("/Volumes/Recovery/projects/r2t/benchmark/like_shen/mouse_tree_compare_assembly-like_shen_all.csv", header=T, sep=',')
dim(mouse)
yeast <- read.csv("/Volumes/Recovery/projects/r2t/benchmark/like_shen/yeast_tree_compare_assembly-like_shen_all.csv", header=T, sep=',')
dim(yeast)
arath <- read.csv("/Volumes/Recovery/projects/r2t/benchmark/like_shen/arath_tree_compare_assembly-like_shen_all.csv", header=T, sep=',')
dim(arath)
mouse$species <- 'M. musculus'
yeast$species <- 'S. cerevisiae'
arath$species <- 'A. thaliana'

data <- rbind(mouse,yeast,arath)

# data$appraoch <- 'r2t'

data <- data[order(data$dist),]
data$inode_dist <- factor(data$dist, levels = c(0,1,2,3,4,5,6)) # for data data
head(data)
```


# Plot all removal distances for coverage levels

```{r}
data <- data[order(data$coverage),]
data$coverage <- factor(data$coverage, levels = c(0.2,0.5,1,5,10,20)) # for canva data
data$dist <- factor(data$dist, levels = c(0,1,2,3,4,5,6)) # for canva data
data_r2t <- data[data$method == 'r2t',]
#data$gcov <- as.character(data$gcov)
p <- ggplot(data_r2t, aes(dist, nrf, shape=coverage, color=dist)) + geom_jitter(aes(color=same_neighbour), height = 0.005, alpha=1)
#p <- ggplot(data_merge, aes(dist_to_remain_neigh, proportions, group=interaction(bin, gcov), shape=gcov)) + geom_line(color='darkgrey')  + geom_point(aes(colour=inode_dist))
p <- p + facet_grid(technology~species)
#p <- p + scale_x_continuous(trans = 'log2')
p <- p + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"), name="Correct neighbour")
#p <- p + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"), name="bin")
p <- p + theme_bw()
#p <- p + theme(legend.position="bottom")
#p <- p + theme(axis.text.x = element_text(), legend.text = element_text(hjust = 1))
p <- p + xlab("Species removed")
p <- p + ylab("Topological similarity to reference")
# pdf("~/projects/r2t/paper/figure2/tree_OGs_likeshen.pdf", width = 8, height = 6, useDingbats=FALSE)
p
```

# Comparison with assembly

```{r}
data_assemblies <- data[which(data$method == 'assembly'),]
head(data_assemblies)
data_tmp <- data_assemblies
```

```{r}
data <- data[order(data$coverage),]
data$coverage <- factor(data$coverage, levels = c(0.2,0.5,1,5,10,20)) # for canva data
data$dist <- factor(data$dist, levels = c(0,1,2,3,4,5,6)) # for canva data
#data$gcov <- as.character(data$gcov)
p <- ggplot(data, aes(dist, rf, shape=coverage)) + geom_jitter(aes(color=method), height = 0.005, alpha=1, size=2)
#p <- ggplot(data_merge, aes(dist_to_remain_neigh, proportions, group=interaction(bin, gcov), shape=gcov)) + geom_line(color='darkgrey')  + geom_point(aes(colour=inode_dist))
p <- p + facet_grid(technology~species)
#p <- p + scale_x_continuous(trans = 'log2')
p <- p + scale_color_manual(values=c('red', "#56B4E9"), name="Approach")
p <- p + scale_shape_manual(values=c(3, 4, 8, 15, 16, 17), name='Coverage')
#p <- p + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"), name="bin")
p <- p + theme_bw()
#p <- p + theme(legend.position="bottom")
#p <- p + theme(axis.text.x = element_text(), legend.text = element_text(hjust = 1))
p <- p + xlab("Species removal level")
p <- p + ylab("Topological similarity to reference")
p <- p + theme(legend.position="bottom")
# pdf("~/projects/r2t/paper/figure2/tree_OGs_likeshen.pdf", width = 8, height = 6, useDingbats=FALSE)
p
ggsave(plot=p,height=5,width=6,dpi=200, filename="~/Desktop/paper/figure3/fig3D.pdf", useDingbats=FALSE)
```



```{r}
data_a <- data_assemblies[
  with(data_assemblies, order(coverage, dist, technology, species)),
]
data_r <- data_r2t[
  with(data_r2t, order(coverage, dist, technology, species)),
]

data_diff <- data.frame(coverage = data_r$coverage, technology = data_r$technology, species = data_r$species, dist = data_r$dist, nrf_diff = data_r$nrf - data_a$nrf, same_neighbour = data_a$same_neighbour)
data_diff$better <- 'r2t better'
data_diff$better[which(data_diff$nrf_diff > 0)] <- 'assembly better'
data_diff$better[which(data_a$nrf==1)] <- 'only r2t'
data_diff$better[which(data_diff$nrf_diff == 0)] <- 'equal'
dim(data_a)
dim(data_r)

p <- ggplot(data_diff, aes(dist, nrf_diff, shape=coverage))   + geom_jitter(aes(color=better), height = 0.005, width=0.4, size =2)
#p <- ggplot(data_merge, aes(dist_to_remain_neigh, proportions, group=interaction(bin, gcov), shape=gcov)) + geom_line(color='darkgrey')  + geom_point(aes(colour=inode_dist))
p <- p + facet_grid(technology~species)
#p <- p + scale_x_continuous(trans = 'log2')
p <- p + scale_color_manual(values=c("red", "#999999",'darkblue' , "#56B4E9"), name="Group")
# p <- p + scale_shape_manual(values=c(12, 21), name="Approach")
p <- p + scale_shape_manual(values=c(3, 4, 8, 15, 16, 17), name='Coverage')

#p <- p + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"), name="bin")
p <- p + theme_bw()
#p <- p + theme(legend.position="bottom")
#p <- p + theme(axis.text.x = element_text(), legend.text = element_text(hjust = 1))
p <- p + xlab("Species removal level")
p <- p + ylab("Difference betwen normalized rf") #\n r2t-ref vs assembly-ref
p <- p + theme(legend.position="bottom")

# pdf("~/projects/r2t/paper/figure3/nrf_dist_read2tree_vs_assembly.pdf", width = 8, height = 6, useDingbats=FALSE)
p
ggsave(plot=p,height=5,width=6,dpi=200, filename="~/Desktop/paper/figure3/fig3C.pdf", useDingbats=FALSE)
```


```{r}
pre_recal <- read.csv('~/Desktop/precision_recall_90.csv')
pre_recal$dist <- factor(pre_recal$dist, levels = c(0,1,2,3,4,5,6)) # for data data
pre_recal$coverage <- factor(pre_recal$coverage, levels = c(0.2,0.5,1,5,10,20)) # for data data

pre_recal$full_species <- 'M. musculus'
pre_recal$full_species[pre_recal$species == 'yeast'] <- 'S. cerevisiae'
pre_recal$full_species[pre_recal$species == 'arath'] <- 'A. thaliana'

to_plot <- pre_recal[which(pre_recal$precision > 0.5),] #pre_recal[which(!(pre_recal$recall== 0 & pre_recal$precision == 0)),]
p <- ggplot(to_plot, aes(recall, precision, shape=coverage)) + geom_point(aes(colour=dist), size=2)
#p <- ggplot(data_merge, aes(dist_to_remain_neigh, proportions, group=interaction(bin, gcov), shape=gcov)) + geom_line(color='darkgrey')  + geom_point(aes(colour=inode_dist))
p <- p + facet_grid(technology~full_species)
# p <- p + scale_y_continuous(trans = 'log2')
# p <- p + scale_color_manual(values=c("red", "darkgreen"), name="Correct neighbour")
p <- p + scale_color_viridis(discrete = TRUE)
p <- p + scale_shape_manual(values=c(3, 4, 8, 15, 16, 17), name='Coverage')
#p <- p + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"), name="bin")
p <- p + theme_bw()
#p <- p + theme(legend.position="bottom")
#p <- p + theme(axis.text.x = element_text(), legend.text = element_text(hjust = 1))
p <- p + xlab("recall")
p <- p + ylab("precision")
# pdf("~/Desktop/paper/figure2/pres_recal-90.pdf", width = 8, height = 6, useDingbats=FALSE)
p
# ggsave(plot=p,height=5,width=6,dpi=200, filename="~/Desktop/paper/figure2/pres_recal-90.pdf", useDingbats=FALSE)

```
